\documentclass[11pt,a4paper]{moderncv}


\usepackage{graphicx} %package to manage images
\graphicspath{ {./images/} }

% moderncv themes
\moderncvtheme[blue]{classic}                 % optional argument are 'blue'
% (default), 'orange', 'red', 'green', 'grey' and 'roman' (for roman fonts, instead of sans serif fonts) \moderncvtheme[orange]{classic}                % idem
\input{moderncv-overrides.tex}

% character encoding
\usepackage[utf8]{inputenc} % Kodowanie UTF8
\usepackage[MeX]{polski} % Wsparcie dla PL

% adjust the page margins
\usepackage[scale=0.8]{geometry}
%\setlength{\hintscolumnwidth}{3cm}						% if you want to change the width of the column with the dates
%\AtBeginDocument{\setlength{\maketitlenamewidth}{6cm}}  % only for the classic
% theme, if you want to change the width of your name placeholder (to leave more space for your address details
%\AtBeginDocument{\recomputelengths}                     % required when changes
% are made to page layout lengths

% personal data
\familyname{\\Jakub Bartczuk}
\email{bartczukkuba@gmail.com}
\social[linkedin][https://pl.linkedin.com/in/jakub-bartczuk-22179189]{linkedin}
\social[github][github.com/lambdaofgod]{github.com/lambdaofgod}
\social[stackoverflow][https://stackoverflow.com/users/2610004/jakub-bartczuk]{stackoverflow}
\social[cross-validated][https://stats.stackexchange.com/users/121270/jakub-bartczuk]{cross-validated}
\social[quora][https://www.quora.com/profile/Kuba-Bartczuk]{quora}
\title{Machine Learning Engineer}               % optional, remove the line if not wanted
\address{}{}    % optional, remove the


\begin{document}
\maketitle

\section{Intro}
\cvitem{}{Data scientist/ML engineer. Solved problems in image processing, sound processing, information retrieval and NLP. Used deep learning, supervised, weakly supervised, unsupervised, transfer and few/zero-shot learning. Mostly machine learning in  real world, low quality/small/noisy data regimes.}

\section{Skills}
\cvitem{General skills}{OOP, big data, functional programming}
\cvitem{Programming Languages}{
\begin{itemize}
    \item Good knowledge: Python, Scala
    \item Basic knowledge: R, Octave/Matlab, Java, Clojure, Haskell, C++
\end{itemize}{}
}
\cvitem{Libraries, frameworks}{numpy, scipy, pandas, pytest}
\cvitem{Deep Learning}{Keras, Tensorflow, PyTorch, MXNet/Gluon, FastAI, Theano}
\cvitem{ML Tools}{scikit-learn, gensim, NLTK, XGBoost, BentoML} 
\cvitem{Cloud Computing}{AWS, GCP}
\cvitem{Other}{scikit-image, Docker, Kubernetes, Git, Spark, Elasticsearch, Linux}

\section{Projects, blog posts, talks}
\cvline{Open Source Projects}{}
\cvline{Creator}{\href{https://github.com/lambdaofgod/findkit}{findkit: A Python library for Content-Based Information Retrieval. Multimodal search using transfer learning and approximate nearest neighbors.
}}
    \cvline{Contributor}{\href{https://github.com/gaetangate/text-summarizer}{text\_summarizer: Extractive Summarization, implementation of paper "Centroid-based Text Summarization through Compositionality of Word Embeddings".
}}

\cvline{Blog post, January 2019}{\href{https://semantive.com/text-summarization-in-python/}{Overview of state of Text Summarization in Python}}

\cvline{Meetup talk, March 2017}{\href{https://github.com/lambdaofgod/lucene4_cookbook_examples/blob/master/notebooks/pydata_geek_week_example.ipynb}{Extending Java/Scala projects with Jupyter-Scala. Talk was given at PyData Wrocław meeting.
}}

\cvline{Course Project}{
\href{https://github.com/lambdaofgod/tda-project}{Topological Data Analysis methods for evaluating nonlinear dimensionality reduction (project for Data Mining course). Uses Javaplex (Java TDA library) from Scala.}}

\newpage

\section{Work Experience}
\cventry{July 2019 - Present}{Machine Learning Engineer}{xBerry}{}{}{
\begin{itemize}
    \item Plagiarism detection engine - text and images. Planning, implementing and deployment
    \item Nudity detection with Computer vision
    \item Scheduling using artificial intelligence
\end{itemize}
}
\cventry{Oct 2017 - Jun 2019}{Data Scientist}{Semantive}{}{}{
\begin{itemize}
 \item Created and deployed search engine for domain specific images using CNNs and approximate nearest neighbours index.
 \item Used extractive text summarization techniques to build service for effective summarization of scientific documents.
 \item Created music search engine using transfer learning on Fourier features (STFT).
 \item Applied Image to Image Translation (Conditional GANs) to image coloring.
 \item Useq Sequence 2 Sequence methods for abstractive text summarization.
 \item Tutored interns and junior data scientists.
 \end{itemize}
}
\cventry{Nov 2016 - Sep 2017}{Big Data Engineer}{IT Kontrakt}{(part-time)}{}
{Worked as a consultant in financial institution.
\begin{itemize}
 \item Developed Scala application with some elements in Spark.
 \item Optimized rule-based classification engine using machine learning (classification algorithms).
\end{itemize}
}

\cventry{Jun 2016 - Sep 2016}{Software Developer - Data Mining}{Neurosoft}{(full-time)}{}
{ Worked on backend of big data application written in Scala and Akka.
}

\cventry{Apr 2016 - Jun 2016}{Scala Developer}{Datarino}{(part-time)}{}
{Worked on backend of Scala application (Akka-Http)}


\cventry{Oct 2015 - May 2016}{Scala Developer}{Helios Ventures Limited}{(part-time)}{}
{Worked on backend of Scala application of social media app (written using Play framework)}
	       
\section{Courses, certificates}
\cvline{Feb 2020}{\href{https://www.coursera.org/account/accomplishments/certificate/YR6KFXVLZE87}{End-to-End Machine Learning with TensorFlow on GCP}}
\cvline{Apr 2019}{\href{https://courses.edx.org/certificates/11126c558a9e4a7697593e15fb591383}{Quantum Machine Learning}}
\cvline{Mar 2019}{\href{https://courses.edx.org/certificates/7c5d40fb30b8456ebe6d774cec098bf5}{Sparse Representations in Image Processing: From Theory to Practice}}
\cvline{Dec 2018}{\href{https://courses.edx.org/certificates/c167c6fca2b847ee9555b4af980ca80e}{Sparse Representations in Signal and Image Processing: Fundamentals}}
\cvline{Dec 2018}{\href{https://www.coursera.org/account/accomplishments/verify/CSQCR26AKASB}{Deep Learning in Computer Vision (Advanced ML Specialization)}}
\cvline{Oct 2018}{\href{https://www.coursera.org/account/accomplishments/verify/SY7WYHK4P65T
}{Natural Language Processing (Advanced ML Specialization)}}
\cvline{Aug 2018}{ \href{https://www.coursera.org/account/accomplishments/certificate/EGYZQHCJ9UAG}{Bayesian Methods for Machine Learning (Advanced ML Specialization)}}
\cvline{Apr 2018}{ \href{https://www.coursera.org/account/accomplishments/certificate/T59FKJT9C429}{Hinton's Neural Networks for Machine Learning}}
\cvline{Mar 2018}{\href{https://udemy-certificate.s3.amazonaws.com/pdf/UC-QIJJ6U91.pdf}{Unsupervised Deep Learning in Python (Udemy)
}}
\cvline{Mar 2018}{\href{https://www.coursera.org/account/accomplishments/specialization/ZNT3DWVS3TFL}{Deep Learning Specialization (Coursera)}}
\cvline{Dec 2017}{AWS Certified Solutions Architect Associate}
\cvline{Sep 2017}{\href{https://www.udemy.com/certificate/UC-8UEXC6U2/}{Data Science: Practical Deep Learning in Theano + TensorFlow (Udemy)}}
\cvline{Jan 2017}{\href{https://courses.cognitiveclass.ai/certificates/user/327568/course/course-v1:BigDataUniversity+BD0221EN+2016}{Spark MLLib}}
\cvline{Dec 2016}{\href{https://courses.cognitiveclass.ai/certificates/user/327568/course/course-v1:BigDataUniversity+ML0101EN+2016_T2}{Machine Learning with Python
}}
\cvline{Apr 2016}{
\href{https://prod-cert-bucket.s3.amazonaws.com/downloads/f864b9d7e7a14efebab0beb26adcd931/Statement.pdf}{Statistical Learning (Stanford Lagunita)}
}
\cvline{Dec 2015}{\href{https://www.youracclaim.com/badges/b1473fc7-21ee-4f46-ad96-c981c78b6d3d/linked_in_profile}{Big Data Spark Foundations}}
\cvline{Sep 2014}{
\href{https://www.coursera.org/maestro/api/certificate/get_certificate?course_id=972303}{Machine Learning (Andrew Ng on Coursera)}
}

\newpage
\section{Education}
\cventry{2019-now}{Data Science MSc}{University of Wrocław}{}{}{}
\cventry{2011 - 2014}
{BS studies in Mathematics}{University of Wrocław}{}{}
{Theoretical Specialization
Relevant courses: Data Mining, Text Mining, Wavelet Analysis. Activities: talks for Theoretical Math Student group (Topology, Quantum Computing)
}

 
\section{Languages}
\cvline{}{Polish (native), English (Full professional proficiency)}



\section{Interests and activities}
\cvline{}{mathematics, science, ancient history, biking, weight lifting}

\closesection{}
\renewcommand{\listitemsymbol}{-}

\end{document}
